# Causal Impact Demonstration

Here we will demonstrate how to use the `CausalImpact` package using
stock market data around the time when Google changed its corporate structure
with the creation of parent company Alphabet.

```{r message = FALSE}
library(CausalImpact)
library(tidyverse)
library(causaldata)
library(zoo)
```

# 1. Data prep and exploration

```{r}
event <- ymd("2015-08-10")
df_google <- google_stock %>%
  filter(Date >= event - 21 & Date <= event + 14) %>%
  mutate(
    Google_Return = Google_Return * 100,
    SP500_Return = SP500_Return * 100
  )
head(df_google)
```

```{r}
ggplot(df_google) +
  geom_line(aes(x = Date, y = Google_Return), color = "red") +
  geom_line(aes(x = Date, y = SP500_Return), color = "blue") +
  geom_vline(aes(xintercept = event), linetype = "dashed") +
  labs(
    title = "Google Stock Price Over Time",
    x = "Date",
    y = "Percent Daily Return"
  )
```

```{r}
df <- read.zoo(df_google)
pre_period <- c(index(df)[1], event)
post_period <- c(event + 1, index(df)[nrow(df)])
head(df)
```

# 2. Run the analysis

```{r}
set.seed(123)
impact <- CausalImpact(df, pre_period, post_period)
summary(impact)
```

| Metric | Value |
| ------ | ----- |
| Actual (Average) | -0.69|
| Actual (Cumulative) | -6.88|
| Predicted (Average) | -1.2 (95% CI: -2.1, -0.35)|
| Predicted (Cumulative) | -12.0 (95% CI: -20.7, -3.46)|
| Absolute Effect (Average) | 0.51 (95% CI: -0.34, 1.4)|
| Absolute Effect (Cumulative) | 5.09 (95% CI: -3.42, 13.8)|
| Relative Effect (Average) | 95% (95% CI: -67%, 88%)|
| Relative Effect (Cumulative) | 95% (95% CI: -67%, 88%)|
| Posterior tail-area probability p-val | 0.114|
| Posterior prob. of a causal effect | 89%|

* Actual (Average): the observed average value of your response variable during the post-intervention period.
* Actual (Cumulative): the observed sum of your response variable during the post-intervention period.
* Predicted (Average): the model's estimated average value of your response variable in the post-intervention period, if the intervention had not occurred (i.e., the counterfactual).
* Predicted (Cumulative): the model's estimated sum of your response variable in the post-intervention period, if the intervention had not occurred.
* Absolute Effect (Average): Actual (Average) - Predicted (Average)
* Absolute Effect (Cumulative): Actual (Cumulative) - Predicted (Cumulative)
* Relative Effect (Average): Absolute (Average) / Predicted (Average)
* Relative Effect (Cumulative): Absolute (Cumulative) / Predicted (Cumulative)
* Posterior tail-area probability p-val: this p-value is a Bayesian analogue to the frequentist p-value. It represents the probability of observing an effect as large as, or larger than, the one estimated, purely by chance, assuming the null hypothesis of no effect is true.
* Posterior prob. of a causal effect: the posterior probability that the actual effect (positive or negative) is non-zero

It appears that there is *some* evidence of a causal effect, but this really
isn't strong enough evidence to confidently claim a causal impact.

```{r}
plot(impact)
```

# 3. Refining the model

```{r}
impact2 <- CausalImpact(
  df,
  pre_period,
  post_period,
  model.args = list(niter = 5000, nseasons = 5)
)
summary(impact2)
```
niter default is 1000 and nseasons default = 1