# TMLE

Objective: ATE of the X-Y effect in the df_uc_source dataframe.

```{r setup, include=FALSE}
library(tidyverse)
library(multibias)
library(tmle)
```

```{r}
df <- df_uc_source |>
  mutate(
    Y = Y_bi,
    X = X_bi
  ) |>
  select(-X_bi, -Y_bi, -X_cont, -Y_cont)

head(df)
```

## With TMLE function

```{r}
tmle(
  Y = df$Y,
  A = df$X,
  W = df |> select(C1, C2, C3, U),
  family = "binomial",
  Q.SL.library = c("SL.glm", "SL.glmnet"),
  g.SL.library = c("SL.glm", "SL.glmnet")
)
# ATE = 0.102, 95% CI = (0.097, 0.108)
# OR = 1.95, 95% CI = (1.89, 2.02)
```

Used here:
* Y: outcome variable
* A: exposure variable
* W: covariates
* family: outcome distribution
* Q.SL.library: SuperLearner library for outcome model
* g.SL.library: SuperLearner library for propensity model

Some other useful parameters not used above:
* Q: an nx2 matrix of predicted outcome probabilities for each observation under each treatment level (column order is A=0, A=1); alternative to using Q.SL.library
* Z: intermediate covariate; for controlled direct effect estimation
* Delta: indicator of missing outcome or treatment


## Step by Step

### 1. Probability of outcome

```{r}
# fit outcome model

Q_mod <- glm(
  Y ~ X + C1 + C2 + C3 + U,
  data = df,
  family = "binomial"
)

# predict in original data

Y_pred_observed <- predict(
  Q_mod,
  type = "response"
)

# predict when exposure = 1

Y_pred_X1 <- predict(
  Q_mod,
  newdata = mutate(df, X = 1),
  type = "response"
)

# predict when exposure = 0

Y_pred_X0 <- predict(
  Q_mod,
  newdata = mutate(df, X = 0),
  type = "response"
)
```

```{r}
# confirm all predictions are between 0 and 1

all(Y_pred_observed >= 0 & Y_pred_observed <= 1) # TRUE
all(Y_pred_X1 >= 0 & Y_pred_X1 <= 1) # TRUE
all(Y_pred_X0 >= 0 & Y_pred_X0 <= 1) # TRUE
```

### 2. Probability of treatment

```{r}
# fit treatment model

g_mod <- glm(
  X ~ C1 + C2 + C3 + U,
  data = df,
  family = "binomial"
)

# inverse probability of treatment

IP_X1_pred <- 1 / predict(
  g_mod,
  type = "response"
)

# negative inverse probabilty of not receiving treatment

IP_X0_pred <- -1 / (1 - predict(
  g_mod,
  type = "response"
))
```

```{r}
# clever covariate

df$cc <- ifelse(df$X == 1, IP_X1_pred, IP_X0_pred)
```

### 3. Fluctuation parameter

```{r}
eps_mod <- glm(
  Y ~ -1 + offset(qlogis(Y_pred_observed)) + cc,
  data = df,
  family = binomial
)

fluctuation_param <- coef(eps_mod)[1]
```

### 4. Effect estimation

```{r}
```

### 5. Standard errors

```{r}
```
